<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  error = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rrq

<!-- badges: start -->
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![R-CMD-check](https://github.com/mrc-ide/rrq/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mrc-ide/rrq/actions/workflows/R-CMD-check.yaml)
[![codecov.io](https://codecov.io/github/mrc-ide/rrq/coverage.svg?branch=master)](https://codecov.io/github/mrc-ide/rrq?branch=master)
[![CodeFactor](https://www.codefactor.io/repository/github/mrc-ide/rrq/badge)](https://www.codefactor.io/repository/github/mrc-ide/rrq)
<!-- badges: end -->

Task queues for R, implemented using Redis.

## Getting started

Create an `rrq_controller` object

```{r}
obj <- rrq::rrq_controller$new("rrq:readme")
```

Submit work to the queue:

```{r}
t <- obj$enqueue(runif(10))
t
```

Query task process:

```{r}
obj$task_status(t)
```

Run tasks on workers in the background

```{r}
rrq::rrq_worker_spawn(obj)
```

Collect task results when complete

```{r}
obj$task_wait(t)
```

Or try and retrieve them regardless of if they are complete

```{r}
obj$task_result(t)
```

Query what workers have done

```{r}
obj$worker_log_tail(n = Inf)
```

For more information, see `vignette("rrq")`

```{r, include = FALSE}
obj$destroy(timeout_worker_stop = 10)
```

## Installation

Install from the mrc-ide package repository:

```r
drat:::add("mrc-ide")
install.packages("rrq")
```

Alternatively, install with `remotes`:

```r
remotes::install_github("mrc-ide/rrq", upgrade = FALSE)
```

## Testing

To test, we need a redis server that can be automatically connected to using the `redux` defaults.  This is satisfied if you have an unauthenticated redis server running on localhost, otherwise you should update the environment variable `REDIS_URL` to point at a redis server.  Do not use a production server, as the package will create and delete a lot of keys.

A suitable redis server can be started using docker with

```
./scripts/redis start
```

(and stopped with `./scripts/redis stop`)

## License

MIT © Imperial College of Science, Technology and Medicine
